Categorizing the trajectory dynamics of replay

### 2019-07-16 Article.tex
Update on Overleaf.

To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics. Each discrete latent “category” is associated with a type of continuous latent dynamic—such as hovering, spatially fragmented or spatially continuous. This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories. In addition, the model can function in 2D, avoiding linearization errors on more complicated environments. We demonstrate the utility of this model on simulated and real data of an animal performing a spatial memory task.

@@

\begin{keywords}
Hippocampus | Replay | State Space | Point Process | Switching
\end{keywords}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{Figures/Figure_1}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{fig:computerNo}
\end{figure}

@@

\section*{Conclusions}

@@

\section*{Conclusions}

@@

%% You can use these special %TC: tags to ignore certain parts of the text.
%TC:ignore
%the command above ignores this section for word count
\onecolumn
\newpage

\section*{Word Counts}
This section is \textit{not} included in the word count. 
\subsection*{Notes on Nature Methods Brief Communication}
\begin{itemize}
\item Abstract: 3 sentences, 70 words.
\item Main text: 3 pages, 2 figures, 1000-1500 words, more figures possible if under 3 pages
\end{itemize}

\subsection*{Statistics on word count}
\detailtexcount
\newpage

>>>>

To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics. Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous. This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories. In addition, the model can function in 2D, avoiding linearization errors on more complicated environments. We demonstrate the utility of this model on simulated and real data of an animal performing a spatial memory task.

@@

\begin{keywords}
Hippocampus | Replay | State Space
\end{keywords}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{figures/Figure_1}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{fig:computerNo}
\end{figure}

@@

\section*{Discussion}

@@

@@

\onecolumn
\newpage

\section*{Materials and Methods}

\subsection*{Filter Derivation}
\begin{enumerate}
\item Use Bayes theorem to define the posterior in terms of a data-derived likelihood and a prior ($ posterior \propto likelihood * prior $)
$$
p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k-1})
$$

\item Use Chapman-Kolmogorov to define prior in terms of previous time step's posterior and state transitions
$$
\begin{align*}
p(x_{k}, I_{k} \mid y_{1:k-1}) &= \sum_{I_{k-1}} \int p(x_{k}, x_{k-1}, I_{k}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
\\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}, y_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
\\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
\\ &= \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}, x_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
\end{align*}$$

where:
$p(x_{k-1}, I_{k-1} \mid y_{1:k-1})$ is the previous posterior,
$Pr(I_{k} \mid I_{k-1}, x_{k-1})$ is the discrete state transition, and $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$ is the continuous state transition

\item Final Filter
$$
p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}, x_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
$$
\end{enumerate}

\subsection*{Smoother Derivation}

\begin{enumerate}
\item
$$
\begin{align*}
p(x_{k}, I_{k} \mid y_{1:T}) &= \sum_{I_{k+1}} \int p(x_{k}, x_{k+1}, I_{k}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
\\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:T}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
\\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:k}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k}, I_{k}, x_{k+1}, I_{k+1} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
\\ &= p(x_{k}, I_{k} \mid y_{1:k}) * \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
\\ &= p(x_{k}, I_{k} \mid y_{1:k}) * \sum_{I_{k+1}} \int \frac{p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}, x_{k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
\end{align*}$$

\item
$$
\begin{align*}
p(x_{k+1}, I_{k+1} \mid y_{1:k}) &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1}, x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
\\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
\\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
\\ &= \sum_{I_{k}} \int p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}, x_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
\end{align*}$$

\end{enumerate}

\newpage
